Project Write-Up - Practical Machine Learning
========================================================
The goal of this project is to build a machine learning algorithm to predict activity quality from activity monitors.

First, we load some libraries along with the datasets and prepare it for analysis / model development

```{r}
library(caret)
library(ggplot2)
library(knitr)

# for reproducibility
set.seed(1337)

# This file assumes we're in same directory as data
# setwd("coursera/predMachLearn")
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")

```

With the data loaded, we can clean it for processing:
```{r}

#clean data (some of these code snippets are modified from the forums):
NAs <- apply(training,2,function(x) {sum(is.na(x))})  #<FORUM SNIPPET>
validData <- training[,which(NAs == 0)] #<FORUM SNIPPET>

trainIndex <- createDataPartition(y = validData$classe, p=0.1,list=FALSE) # pared this down so it won't take forever to knit...
trainData <- validData[trainIndex,]

# discards unuseful predictors 
removeIndex <- grep("timestamp|X|user_name|new_window",names(trainData)) # <FORUM SNIPPET>
trainData <- trainData[,-removeIndex]

# I'll bet we can get pretty accurate using just the numeric sensor data
num_only <- sapply(trainData, is.numeric)
# that would blow away the 'classe' column so I need to add that back in
# before filtering:
num_only[87]=TRUE

trainData <- trainData[,num_only]

```

With the data cleaned, we can proceed to train the model:
```{r}

tr = trainControl(method = "cv", number = 4) # < FORUM CODE SNIPPET, this really sped things up and improved memory use >

# make model

modFit <- train(classe ~.,data = trainData,method="rf",preProcess=c("center","scale","pca"), trControl=tr)

# check our work:
modFit
```


Not sure what to expect for error, so need to cross validate on the test set.

```{r}
# try on the Cross Validation set:

xValTest <- training[-trainIndex,]
xValPred <- predict(modFit,xValTest)
predRight <- xValPred == training[-trainIndex,]$classe
accuracy <- sum(predRight)/length(predRight)

accuracy
# Accuracy was > 97% for training with 40% of the data

```

```{r}
# Finally, do the test set:

testPred <- predict(modFit,testing)

# and load submit function:

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(testPred)

# Files are loaded manually. Model classified 17/20 correctly so I'm happy with that.

```


